### 2016-07-25 Manuscript/1.tex
Kempes edits to v2

%\usepackage{scicite}
%\usepackage{times}
%\usepackage{graphicx}
%\usepackage{verbatim}
%\usepackage{subfigure}
%\usepackage{tikz}
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{array}
%\usepackage{tabularx}
%\usepackage{multirow}
%\usepackage{multicol}
%\usepackage{multibox}
%\usepackage{rotating}
%\usepackage{layout}
%\usepackage{epstopdf}
%\usepackage{afterpage}
%\usepackage[left]{lineno}

%% PNAStmpl.tex
%% Template file to use for PNAS articles prepared in LaTeX
%% Version: Apr 14, 2008

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% BASIC CLASS FILE

\documentclass{pnastwo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% OPTIONAL GRAPHICS STYLE FILE

\usepackage[pdftex]{graphicx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% OPTIONAL POSTSCRIPT FONT FILES

\usepackage{pnastwof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% ADDITIONAL OPTIONAL STYLE FILES

%\usepackage{amssymb,amsfonts,amsmath}
\usepackage{amsmath}

\usepackage[font=large]{caption}
\usepackage{subfigure}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{array}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{color}
%\usepackage{multibox}
\usepackage{rotating}
%\usepackage{color}
%\usepackage{lineno}
\usepackage[left]{lineno}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% OPTIONAL MACRO FILES

\bibliographystyle{pnas}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\sid}[1]{\textcolor{red}{\bf [#1]}}

%% Don't type in anything in the following section:
%%%%%%%%%%%%
%% For PNAS Only:
\contributor{Submitted to Proceedings
of the National Academy of Sciences of the United States of America}
\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
\copyrightyear{2011}
\issuedate{Issue Date}
\volume{Volume}
\issuenumber{Issue Number}
%%%%%%%%%%%%

\begin{document}

\linenumbers
\setlength\linenumbersep{3pt}

\title{Ecological and evolutionary implications of starvation and body size}

%\title{Survival of the Stable: Dynamic Consequence of Starvation and Reproduction}

%\title{Size Constraints on Starvation/Reproduction Dynamics}

\author
{Justin D. Yeakel\affil{1}{School of Natural Science, University of California Merced, Merced, CA} \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{To whom correspondence should be addressed: jdyeakel@gmail.com}, Christopher P. Kempes \affil{2}{The Santa Fe Institute, Santa Fe, NM}, \and Sidney Redner \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{3}{Department of Physics, Boston University, Boston MA}
}

\contributor{Submitted to Proceedings of the National Academy of Sciences
of the United States of America}

\maketitle

\begin{article}

\begin{abstract}
This is the abstract.  \emph{ No it isn't.  It's merely a placeholder.}
\end{abstract}
\keywords{foraging | starvation|reproduction}

\noindent {\bf Introduction} \\ \nonumber
The behavioral ecology of most, if not all, organisms is influenced by the
energetic state of individuals, which directly influences how organisms
invest reserves in uncertain environments.  Such behaviors are generally
manifested as trade-offs between investing in somatic maintenance and growth
or allocating energy towards reproduction
\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}. 
The timing of these behaviors responds to selective pressure, as the choice
of the investment impacts future fitness!\cite{Mangel:1988uaa}.
The influence of resource limitation on an organism's ability to maintain its
nutritional stores may lead to repeated delays or shifts in reproduction over
the course of an organism's life.

The life history of most species is typically comprised of (a) somatic growth
and maintenance and (b) reproduction.  The balance between these two
activities is often conditioned on resource
availability~\cite{Morris:1987eo}.  For example, reindeer invest less in
calves born after harsh winters (when the mother's energetic state is
depleted) than in calves born after moderate winters \cite{Tveraa:2003fq}.
Many bird species invest differently in broods during periods of resource
scarcity compared to normal periods~\cite{Daan:1988va,Jacot:2009dv},
sometimes delaying or even foregoing reproduction for a breeding
season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.  Even freshwater
and marine zooplankton have been observed to avoid reproduction under
nutritional stress \cite{Threlkeld:1976ih}, and those that do reproduce have
lower survival rates~\cite{Kirk:1997cc}.  Organisms may also separate
maintenance and growth from reproduction over space and time: many salmonids,
birds, and some mammals return to migratory breeding grounds to reproduce
after one or multiple seasons in resource-rich environments where they
accumulate nutritional
reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.

Physiological mechanisms also play an important role in regulating
reproductive expenditures during periods of resource limitation.  Diverse
mammals (47 species in 10 families) exhibit delayed implantation, whereby
females postpone fetal development (blastocyst implantation) until times
where nutritional reserves can be
accumulated~\cite{Mead:1989dt,Sandell:1990kw}.  Many other many species
(including humans) suffer irregular menstrual cycling and higher abortion
rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.
In the extreme case of unicellular organisms, nutrition is unavoidably linked
to reproduction because the nutritional state of the cell regulates all
aspects of the cell cycle \cite{Glazier:2009hq}.  The existence of so many
independently evolved mechanisms across such a diverse suite of organisms
highlights the importance and universality of the fundamental tradeoff
between somatic and reproductive investment.  However the dynamic
implications of these constraints are unknown.

Though straightforward conceptually, incorporating the energetic dynamics of
individuals~\cite{Kooi2000} into a population-level
framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical
obstacles~\cite{Diekmann:2010da}).  An alternative approach involves modeling
the macroscale relations that guide somatic versus reproductive investment in
a consumer-resource system.  For example, macroscale Lotka-Volterra models
assume that the growth rate of the consumer population depends on resource
density, thus \emph{implicitly} incorporating the requirement of resource
availability for reproduction \cite{murdoch:2003}.

In this work, we adopt an alternative approach in which resource limitation
and the subsequent effect of starvation is accounted for \emph{explicitly}.
Namely, only individuals with sufficient energetic reserves can reproduce.
Such a constraint leads to reproductive time lags due to some members of the
population starving and then recovering.  Additonally, we incorporate the
idea that reproduction is strongly constrained
allometrically~\cite{Kempes:2012hy}, and is not generally linearly related to
resource density.  As we shall show, these constraints influence the ensuing
population dynamics in dramatic ways.
\\

\noindent {\bf Nutritional-state-structured model (NSM)}

We begin by defining a minimal Nutritional-State-structured population Model
(NSM), where the consumer population is divided into two energetic states:
(a) an energetically replete (full) state $F$, where the consumer reproduces
at a constant rate $\lambda$, and (b) an energetically deficient (hungry)
state $H$, where the consumer does not reproduce but dies at rate $\mu$.  The
underlying resource $R$ evolves by logistic growth with an intrinsic growth
rate $\alpha$ and a carrying capacity equal to one.  Consumers transition
from the full state $F$ to the hungry state $H$ by starvation at rate
$\sigma$ and also in proportion to the absence of resources $(1-R)$.
Conversely, consumers recover from state $H$ to state $F$ at rate $\rho$ and
in proportion to $R$.  Resources are also eaten by the consumers---at rate
$\rho$ by hungry consumers and at rate $\beta<\rho$ by full consumers.  This
inequality accounts for hungry consumers requiring more resources to rebuild
body weight.

In the mean-field approximation, in which the consumers and resources are
perfectly mixed, their densities evolve according to the rate equations
\begin{eqnarray} 
\label{eq:system}
\begin{split}
\dot{F} &= \lambda F + \rho RH - \sigma (1-R)F,  \\
\dot{H} &= \sigma (1-R)F - \rho RH - \mu H,  \\
\dot{R} &= \alpha R(1-R) - R(\rho H+ \beta F).
\end{split}
\end{eqnarray}
Notice that the total consumer density $F+H$ evolves according to
$\dot{F}+\dot{H}=\lambda F-\mu H$.  This resmebles the equation of motion for
the predator density in the classic Lotka-Volterra model, except that the
resource density does not appear in the growth term.  As discussed above, the
attributes of reproduction and mortality have been explicitly apportioned to
the full and hungry consumers, respectively, so that the growth in the total
density is decoupled from the resource density.

Equation~\eqref{eq:system} has three fixed points: two trivial fixed points
at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal
fixed point at
\begin{eqnarray}
\label{eq:ss}
\begin{split}
F^* &= \frac{\alpha  \lambda  \mu  (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  \beta)}, \\
H^* &= \frac{\alpha  \lambda ^2 (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  \beta)}, \\
R^* &= \frac{\mu  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma }.
\end{split}
\end{eqnarray}
If this internal fixed point, which is unique, is stable, it will be the
global attractor for all population trajectories for any initial condition
where the resource and consumer densities are both non zero.  The stability
of this fixed point is determined by the Jacobian Matrix $\bf J$, where each
matrix element $J_{ij}$ equals $\partial{\dot X_i}/\partial{X_j}$ when
evaluated at the internal fixed point, and $\mathbf{X}$ is the vector
$(F,H,R)$.  If the parameters in Eq.~\eqref{eq:system} are such that the real
part of the largest eigenvalue of $\mathbf{J}$ is negative, then the system
is stable with respect to small perturbations from the fixed point.

From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the
reproduction rate $\lambda$ must be less than the starvation rate $\sigma$,
so that $R^*$ is positive.  Moreover, when the resource density $R=0$, the
rate equation for $F$ gives exponential growth for $\lambda>\sigma$.  The
condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation that
demarcates the physical and unphysical regimes.  The biological implication
of the constraint $\lambda<\sigma$ is simple---the rate at which a
macroscopic organism loses mass due to lack of resources is generally much
faster than the rate of reproduction.  As we will discuss below, this
inequality is a natural consequence of allometric
constraints~\cite{Kempes:2012hy} for organisms within empirically observed
body size ranges (Fig.~\ref{fig:gvs}).

In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may
either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).  In
continuous-time systems, a limit cycle arises when a pair of complex
conjugate eigenvalues crosses the imaginary axis to attain positive real
parts~\cite{GuckHolmes}.  This Hopf bifurcation is defined by
${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is
composed of the coefficients of the characteristic polynomial of the Jacobian
matrix~\cite{Gross:2004p2428}.  As the system parameters are tuned to be
within the stable regime but close to the Hopf bifurcation, the amplitude of
the transient but decaying cycles become large.  Given that ecological
systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of
transient cycles, even though they decay with time, can increase the
extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.  Thus
the distance of a system from the Hopf bifurcation provides a measure of its
persistence.

When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the
consumers are driven to the hungry non-reproductive state.  Because
reproduction is inhibited, there is a low steady-state consumer density and a
high steady-state resource density.  However, if $\sigma/\lambda$ approaches
one, the population is overloaded with energetically-replete (reproducing)
individuals, thereby promoting oscillations between the consumer and resource
densities (Fig.~\ref{fig:fp}).

Whereas the consumer growth rate $\lambda$ defines an absolute bound of
biological feasibility---the TC bifurcation---the starvation rate $\sigma$
determines the sensitivity of the consumer population to changes in resource
density.  When $\sigma\gg\lambda$, the steady-state population density is
small, thereby increasing the risk of stochastic extinction.  Conversely, as
$\sigma$ decreases, the system will ultimately be poised either near the TC
or the Hopf bifurcation (Fig.~\ref{fig:fp}).  If the recovery rate $\rho$ is
sufficiently small, the TC bifurcation is reached and the resource eventually
is eliminated.  If $\rho$ exceeds a threshold value, cyclic dynamics will
develop as the Hopf bifurcation is approached.

%In the case of the 2-stage consumer model, the Jacobian evaluated at the internal steady state (denoted by $|_*$) is written
%
%\begin{equation}
%\mathbf{J}|_* =
%%\left(
%%\begin{array}{ccc}
%% -F^* m+(1-R^*) a -R^* a -H^* \rho  & -R^* \rho  & -m R^* \\
%% -H^* \rho -F^* \sigma  & -\mu -R^* \rho  & (1-R^*) \sigma  \\
%% H^* \rho +F^* \sigma  & R^* \rho  & \lambda -(1-R^*) \sigma  \\
%%\end{array}
%%\right).
%\left(
%\begin{array}{ccc}
% -\frac{\lambda  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\mu  \rho  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
% \frac{\lambda  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\mu  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
% -\frac{m \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } & -\frac{\mu  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } \\
%\end{array}
%\right).
%\end{equation}

%CHRIS - we need to define beta in this section...

\vspace{2mm}
\noindent {\bf Role of allometry} \\ 
%[Link Allometry stuff to our model - what does it provide to the story]
%[Introduce biological and linked constraints]
%[Allometries capture vast amounts of diversity via a single parameter --- body size]
%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
The parameters in the NSM cannot be freely navigated in biologically-reasonable settings, and our first challenge is to constrain the covariation of rates in a principled and biologically meaningful manner. % we have not described the realistic regimes occupied by organisms where the
Allometric scaling relationships highlight common constraints and average trends across large ranges in body size and species diversity. Many of these relationships can be derived from a small set of assumptions and below we describe a framework for the covariation of timescales and rates across the range of mammals for each of the key parameters of our model (cf. \cite{Yodzis:1992hg}). We are able to define the regime of dynamics occupied by the entire class of mammals along with the key differences between the largest and smallest mammals.

%Second paragraph
Nearly all of the rates described in the NSM are to some extent governed by consumer metabolism, and thus can be estimated based on known allometric constraints. 
The scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ is well documented \cite{West:2002it} and plays a central role in a variety of scaling relationships.
Organismal metabolic rate $B$ is known to scale as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to be $3/4$ for metazoans, and varies in unicellular species between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria \cite{DeLong:2010dy}. Several efforts have shown how a partitioning of this metabolic rate between growth and maintenance purposes can be used to derive a general equation for the growth trajectories and growth rates of organisms ranging from bacteria to metazoans \cite{Kempes:2012hy}. More specifically, the interspecific trends in growth rate can be approximated by $\lambda = \lambda_0 M^{\eta-1}$. This relationship is derived from the simple balance
\begin{eqnarray}
\label{balance}
B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m
\end{eqnarray}
 [a and b notation --- these parameters are easily measured bioenergetic parameters which are often approximately invariant across organisms of vastly different size. Our notation seeks to illustrate that the allometric model fundamentally depends on a small number of free parameters.]
 where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is the metabolic rate to support an existing unit of mass, and $m$ is the mass at any point in development. It is useful to explicitly write this balance because it can also be modified to understand the rates of both starvation and recovery from starvation. [Spell out the connection to nutritional state more explicitly] [As we will see it is possible to derive both sigma and rho from this balance]

For the rate of starvation, we make the simple assumption that an organism must meet its maintenance requirements using digested mass as the sole energy source. This assumption implies the simple metabolic balance 
\begin{eqnarray}
\frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
\end{eqnarray}
where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing body mass which may differ from  $E_{m}$, the energy required to synthesis a unit of biomass. Give the adult mass, $M$, of an organism this energy balance prescribes the mass trajectory of a starving organism:
\begin{eqnarray}
m\left(t\right)=Me^{-B_{m}t/E_{m}^{\prime}}.
\end{eqnarray}
Considering that only certain tissues can be digested for energy, for example the brain cannot be degraded to fuel metabolism, we define the rate for starvation and death by the timescales required to reach specific fractions of normal adult mass. We define $m_{starve}=\epsilon M$ where it could be the case that organisms have a systematic size-dependent requirement for essential tissues, such as the minimal bone or brain mass. For example, considering the observation that body fat in mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$, and assuming that once this mass is fully digested the organism begins to starve, would imply that $\epsilon=1-f_{0}M^{\gamma}/M$. Taken together the time scale for starvation is given by
\begin{equation}\label{eq:sigma}
t_{\sigma}=-\frac{E_{m}\log\left(\epsilon\right)}{B_{m}}.
\end{equation}
The starvation rate is $\sigma=1/t_{\sigma}$, which implies that $\sigma$ is independent of adult mass if $\epsilon$ is a constant, and if $\epsilon$ does scale with mass, then $\sigma$ will have a factor of $1/\log\left(1-f_{0}M^{\gamma}/M\right)$. In either case $\sigma$ does not have a simple scaling with $\lambda$ which is important for the dynamics that we later discuss.

The time to death should follow a similar relationship, but defined by a lower fraction of adult mass, $m_{death}=\epsilon^{\prime} M$. Consider, for example, that an organism dies once it has digested all fat and muscle tissues, and that muscle tissue scales with body mass according to $M_{mm}=mm_{0}M^{\zeta}$, then $\epsilon^{\prime}=1-\left(f_{0}M^{\gamma}+mm_{0}M^{\zeta}\right)/M$. Muscle mass has been shown to be roughly proportional to body mass \cite{muscle} in mammals and thus  $\epsilon^{\prime}$ is effectively $\epsilon$ minus a constant. Thus
\begin{eqnarray}
t_{\mu}=-\frac{E_{m}\log\left(\epsilon^{\prime}\right)}{B_{m}}
\end{eqnarray}
and $\mu=1/t_{\mu}$.

%CHRIS - MAYBE WE CUT THIS OUT SINCE WE STICK TO MAMMALS IN THIS PAPER
%It should be noted that we have thus far used mammalian allometry to describe the size-based relationships for growth, starvation, and death. However, our presentation is general, and other functional forms for $\epsilon$, for example, could be determined for other classes of organisms. Considering bacteria, we might expect that starvation or death is defined by the complete digestion of proteins, and in Table \ref{parameter-values} we provide all parameter values for bacteria which we later use as a comparison in our analysis.

The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues tissue from the starving state to the full state.
We again use the balance given in Equation \ref{balance} to find the timescale to return to the mature mass from a given reduced starvation mass. The general solution to Equation \ref{balance} is given by
\begin{eqnarray}
m\left(t\right)=c\left[1-\left(1-\frac{b}{a}m_{0}^{1-\eta}\right)e^{-b\left(1-\eta\right)t}\right]^{1/\left(1-\eta\right)}%\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}
\end{eqnarray}
with $a=B_{0}/E_{m}$, $b=B_{m}/E_{m}$, and $c=(a/b)^{1/(\eta-1)}$. We are then interested in the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, which has the final form of 
\begin{eqnarray}
t_{\rho}=\frac{\log \left(1-\left(cM \right)^{1-\eta }\right)-\log \left(1-\left(c\epsilon M \right)^{1-\eta }\right)}{(\eta -1) b}.
\end{eqnarray}
%\left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}
%where $c=(a/b)^{1/(\eta-1)}$.
Although these rate equations are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from $M\approx1$ gram (the Etruscan shrew \emph{Suncus etruscus}) to $M\approx10^7$ grams (the late Eocene to early Miocene Indricotheriinae).
Investigating other classes of organisms requires only substituting the energetic and scale parameters shown in Table \ref{param}.
Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species. 
%, though this is beyond the current scope of our analysis. %, and we henceforth focus our examination on mammalian species.

\vspace{2mm}
\noindent {\bf The stabilizing effects of allometric constraints} \\ \noindent

%Bifurcation space as a function of mass
As the allometric derivations of NSM rate laws reveal, $\sigma$ and $\rho$ are not independent parameters, and the bifurcation space shown in Fig. \ref{fig:fp} cannot be freely navigated if assuming biologically reasonable parameterizations.
Given the parameterization for terrestrial endotherms shown in Table \ref{param} with mass $M$ as a free parameter, we show that $\sigma$ and $\rho$ are constrained to a small window of potential values (Fig. \ref{fig:hopf}), thus confining dynamics to the steady state regime for all realized body size classes.
%, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
Moreover, for larger $M$, the distance to the Hopf bifurcation increases, while uncertainty in allometric parameters (20\% variation around the mean; Fig. \ref{fig:hopf}) results in little qualitative difference.
This suggests that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than mammals with larger body size. %, and
%2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.

Allometric constraints have been invoked to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, such that period $\propto M^{0.25}$, however this relationship seems to hold only for some species \cite{Hendriks:2012fc} and competing explanations exist \cite{Kendall:1999iy,Hogstedt:2005cr}.
%It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic \cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
Statistically significant support for the existence of population cycles among mammals is predominantly based on time-series for smaller bodied mammals \cite{Kendall:1998hl}, though we acknowledge that longer generation times precludes similar quality data for larger organisms.
%, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
We thus obtain a specific prediction from our model: population cycles should be less common for larger species and more common for smaller species, particularly in environments where resources are limiting. %, emphasizing the role of starvation on dynamics.

%Extinction to the left and the right
Higher rates of starvation result in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality, however it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}).
We examine the competing effects of cyclic dynamics vs. changes in steady state density on extinction risk as a function of the ratio $\sigma/\rho$.
We computed the probability of extinction, where extinction is defined as $H(t)+F(t)=10$ at any instant across all values of $10^2<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq. \ref{eq:system} for an organism of $M=100$ grams, assuming random initial conditions around the steady state (Eq. \ref{eq:ss}). %random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$.
By allowing the rate of starvation to vary, we assessed extinction risk across a range of values of the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a horizontal cross-section of Fig. \ref{fig:fp}.
As expected, higher rates of extinction correlated with both low and high values of $\sigma/\rho$; for low values the higher extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig. \ref{fig:ext}).
For large values of $\sigma/\rho$, higher extinction risk is due to the decrease in the steady state consumer population density.
This interplay creates an `extinction refuge' as shown in Fig. \ref{fig:ext}, such that for a relatively constrained range of $\sigma/\rho$, extinction probabilities are minimized.

As has been described, the $\sigma$ vs. $\rho$ space cannot be freely traversed, such that not all values of $\sigma/\rho$ are biologically feasible.
We observe that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around energetic parameter means) fall within the extinction refuge, such that they are close enough to the Hopf bifurcation to avoid low steady state densities, though far enough away to avoid large-amplitude transient cycles.
The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions driving observed starvation and recovery rates within populations.
Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
Nevertheless, our finding that allometrically-determined energetic rates place the system within this low extinction probability region suggests that the NSM system provides general insight to a phenomena that may both drive -- and constrain -- natural animal populations.

%Results are insensitive to alpha but possibly sensitive to beta (eff parameter)

%NOTES 7/8/16
%Calder has a hypothesis that period ~ Mass^(1/4).
%Can we get "transient period" from Jacobian?
%Lots of food web lit on stabilizing effects of body size! Brose, Petchey for instance. Good for discussion.

\vspace{2mm}
\noindent {\bf Dynamic and energetic barriers to body size} \\ \nonumber
Metabolite transportation constraints are widely thought to place strict boundaries on biological scaling \cite{Brown:1993p708,West:1997cg,Brown:2004wq}, leading to specific predictions on the minimum possible body size for organisms \cite{West:2002ud}.
Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
The \emph{fasting endurance hypothesis} contends that larger body size, with lower metabolic rates and able to hold more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability \cite{Millar:1990p923}.
Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's Rule) \cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper-bound of ca. $10^7$ grams \cite{Alroy:1998p1594}, the value of which may arise from higher extinction risk for large taxa over evolutionary timescales \cite{Clauset:2009fh}.
These trends are thought to be driven by a combination of climate change and niche availability \cite{Saarinen:2014br}, however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, has not been explored, and we contend that the NSM provides a suitable framework to explore these issues.

%Minimal mass bound
The NSM correctly predicts that species with smaller masses have larger steady state population densities, however we observe that there is a sharp asymptote in both steady state densities as well as $\sigma/\rho$ at $M \approx 0.3$ grams  (Fig. \ref{fig:mass}a,b).
Observation of the rates of starvation and recovery explain why: as mass decreases, the rate of starvation increases, while the rate of recovery declines super-exponentially.
This decline in $\rho$ occurs when body fat percentage is $1 - 1/(cM) \approx 2\%$, whereupon consumers have no eligible route out of starvation.
Compellingly, this dynamic bound determined by the rate of energetic recovery is close to the minimum observed mammalian body size of ca. 1.3-2.5 grams (Fig. \ref{fig:mass}b,c), a range that occurs as the recovery rate begins its decline.
In addition to known transport limitations \cite{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.

%Maximum mass bound
Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
Instead, we examine a potential upper bound to body mass by assessing population invasibility with respect to a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-0.5,0.5]$, thus altering rates of starvation, recovery, and maintenance $\beta$.
Although there is not an internal fixed point where both residents and invaders coexist (except for the trivial state $\chi=0$), we can assess invasibility as a function of organismal mass by determining which consumer steady state is larger over $\chi$.
We find that for $1\leq M<10^6$g, having additional body fat ($\chi > 0$) results in a higher steady state density for the invader population ($H^{\prime *}+F^{\prime *}>H^*+F^*$), such that it has an intrinsic advantage over the resident population.
For $M>10^6$, however, there is an increasing range of $\chi < 0$ such that leaner individuals have the advantage, and this is due to the changing covariance between energetic rates as a function of modified energetic reserves.

The observed switch in invasibility as a function of $\chi$ at $M_{\rm opt}\approx10^6$ thus serves as an attractor, where over evolutionary time the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
Moreover, $M_{\rm opt}$, which is entirely determined by the population-level consequences of energetic constraints is within an order of magnitude as that observed in the North American mammalian fossil record \cite{Alroy:1998p1594} and as that predicted from an evolutionary model of body size evolution \cite{Clauset:2009fh}.
While the state of the environment, as well as the competitive landscape, will determine whether specific body size sizes are selected for or against \cite{Saarinen:2014br}, we suggest that the starvation dynamic proposed here may supply the fundamental momentum fueling the evolution of larger body size among terrestrial mammals.
%and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.

%Closing
The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations \cite{Stearns:1989ip}.
The NSM is a minimal and general model that incorporates the dynamics of starvation that are expected to occur in resource limited environments.
By incorporating allometric relationships between the rates in the NSM, we find
\emph{i}) different organismal masses are more or less prone to different population dynamic regimes,
\emph{ii}) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and
\emph{iii}) the dynamic consequences of these rates may place additional barriers on the evolution of minimum and maximum body size.
We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.
Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation.

%In accordance with the fasting endurance hypothesis, 
%
%
%Such energetic advantages, coupled with the 
%Ove and tends to be selected for over evolutionary time.
%The NSM is in accordance with the XXX, where species with larger masses are less likely to evince cyclic dynamics.
%
%
%
%
%We offer an additional mechanism for the minimum and maximum limitation on mammalian body size operating on ecological timescales.
%Such a mechanism would 
%
%
%
%
%Although these physiological barriers are indeed of primary importance, our minimal model of starvation dynamics suggests that there may exist additional constraints due to dynamic and evolutionary effects at small and large body sizes, respectively.
%
%
%

%Looking Forward

%
%
%%Starving Random Walk
%In our modeling, foragers look for food by wandering in this changing
%environment.  If such a search is successful, the forager is satiated and it
%can engage in the essential activity of reproduction.  However, if the search
%is unsuccessful for a sufficiently long period, the forager ``starves''.
%Such a forager can do nothing else but forage, until it either finds food and
%again becomes satiated or it dies when it goes too long without finding
%nourishment.  These rules are reminiscent of the ``starving random walk''
%model, where a single random walk can take $\mathcal{S}$ steps without
%encountering food before starving to death.  Moreover, the resource does not
%regenerate, so that the forager ultimately starves to death.  For this
%idealized model, it was found that the average lifetime of the forager scales
%algebraically with $\mathcal{S}$ in $d\!=\!1$ and $d\!=\!2$ dimensions, and
%as $\exp(-A\mathcal{S}^\omega)$ for $d>2$.  Here the exponent
%$\omega\approx \frac{1}{2}$ for $d=3$, while $\omega\to 1$ only as
%$d\to \infty$, with the latter behavior corresponding to the mean-field
%limit.  As we will discuss, regeneration of the resource, together with the
%behavioral change between starving and satiated foragers leads to still much
%richer dynamical behavior.

%The reproductive tradeoff as a strategy
%

%
%
% \begin{figure}[h]
% 	\centering
% 	\includegraphics[width=0.5\textwidth]{fig_resvuln.pdf}
% 	\caption{
% 	Probability that the resource value is less than threshold = 0.05
% 	}
% 	\label{SN}
% \end{figure}
%
%
%
% \begin{figure}[h]
% 	\centering
% 	\includegraphics[width=0.5\textwidth]{fig_Competition.pdf}
% 	\caption{
% 	Hopf bifurcation
% 	}
% 	\label{SN}
% \end{figure}

\bibliography{aa_starving}

\begin{acknowledgments}
C.P.K was supported by a Trump Fellowship from the American League of Conservatives.
\end{acknowledgments}

%\onecolumn
%\let\cleardoublepage\clearpage

%FIGURE 1

\end{article}

\begin{figure*}
\centering
\includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
\caption{ % 58 words
}
\label{fig:growth}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.5\textwidth]{fig_GvS.pdf}
\caption{ % 58 words
}
\label{fig:gvs}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.5\textwidth]{fig_FixedPoint.pdf}
\caption{ % 58 words
}
\label{fig:fp}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.5\textwidth]{fig_DataHopf.pdf}
\caption{ % 58 words
}
\label{fig:hopf}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric.pdf}
\caption{ % 58 words
}
\label{fig:ext}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.45\textwidth]{fig_FPAllometric.pdf}
\caption{ % 58 words
}
\label{fig:mass}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=1\textwidth]{fig_Invasion.pdf}
\caption{ % 58 words
}
\label{fig:invasion}
\end{figure*}

% 
% 
%	 \begin{figure}[h]
%		 \centering
%		 \includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
%		 \caption{
%		 }
%		 \label{growth-diagram}
%	 \end{figure}
%	 
%	 \begin{figure}[h]
%		 \centering
%		 \includegraphics[width=0.25\textwidth]{fig_GvS.pdf}
%		 \caption{
%		 }
%		 \label{GvS}
%	 \end{figure}
%
%
%	\begin{figure}[h]
%	 	\centering
%	 	\includegraphics[width=0.25\textwidth]{fig_FixedPoint.pdf}
%	 	\caption{
%	 	}
%		\label{Hopfb}
%	\end{figure}
%
%
%	\begin{figure}[h]
%		\centering
%		\includegraphics[width=0.8\textwidth]{fig_DataHopf.pdf}
%		\caption{
%		}
%		\label{DataHopf}
%	\end{figure}
%	 
% 
%	\begin{figure}[h]
%		\centering
%	 	\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometric.pdf}
%	 	\caption{
%	 	}
%	 	\label{Ext}
%	\end{figure}
%	
%	\begin{figure}[h]
%	 	\centering
%	 	\includegraphics[width=0.5\textwidth]{fig_FPAllometric.pdf}
%	 	\caption{
%	 	}
%	 	\label{Asymp}
%	\end{figure}
%
\clearpage

\begin{table}[h]
\caption{Parameter Values For Various Classes of Organisms}
\label{param}
    \begin{center}
    \small
     \begin{tabular}{| p{1.2cm}| p{3.2cm} | p{2.6cm} | p{3.2cm} | }
     \hline
     & {\bf Mammals} & {\bf Unicellular Eukaryotes} & {\bf Bacteria} \\
     \hline
   $\eta$ & $3/4$ & & $1.70$ \\ 
   $E_{m}$ & $10695$ (J gram$^{-1}$) & & $10695$ (J gram$^{-1}$) \\ 
   $E_{m}^{\prime}$ & $\approx E_{m}$ & & $\approx E_{m}$ \\ 
   $B_{0}$ & $0.019$ (W gram$^{-\alpha}$) & & $1.96\times10^{17}$ \\
   $B_{m}$ & $0.025$ (W gram$^{-1}$)   & & $0.025$ (W gram$^{-1}$)\\
   $a$ & $1.78\times10^{-6}$ & & $1.83\times10^{13}$ \\ 
   $b$ & $2.29\times10^{-6}$ & & $2.29\times10^{-6}$ \\  
   $\eta-1$ & $-0.21$ & & $0.73$ \\ 
   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) & & $56493$ \\ 
   $\gamma$ & $1.19$ & & $0.68$ \\ 
   $f_{0}$ & $0.02$ & & $1.30\times10^{-5}$ \\ 
   $\zeta$ & $1.01$ & & \\ 
   $mm_{0}$ & $0.32$ & & \\

   \hline
    \end{tabular}
    \end{center}
   \end{table}

\end{document}